Oregon 2013 Medication for Asthma Inhalers

### 2017-07-13 r_scripts/analysis/oregon_medication.Rmd
Update oregon_medication.Rmd

## Time-Stratified Case-Crossover (Population-Weighted PM~2.5~ Assigned by Zip Code)
*referents same day of week within fire season May - September*

@@

## Time stratified for time series (Population-Weighted PM~2.5~ Assigned by County)

@@

@@

## County-level time series

### County Time series plots of rate

print(plot_county)
print(plot_county_line)
# save figure
ggsave("../plot_new/ndc_county_rate_ts.pdf", plot = plot_county, 
       width = 12, height = 8, units = "in")
ggsave("../plot_new/ndc_county_rate_ts_line.pdf", plot = plot_county_line, 
       width = 12, height = 8, units = "in")

plot(plot_health_smk)
plot(plot_county_health_smk)

### Standard Regression Model (Relative Risk)

1. Fixed-Quasipoisson
2. Fixed-Poisson
3. Mixed-Poisson

```{r regression model q, message=FALSE, echo=FALSE, eval=FALSE, warning=FALSE}
# create empty dataframe
rr_df <- data.frame(matrix(nrow = 3, ncol = 5))
colnames(rr_df) <- c("method", "outcome", "rel_risk", "lower_95", "upper_95")

# methods
rr_df[,1] <- c("Fixed-Quasipoisson", "Fixed-Poisson", "Mixed-Poisson")

# names 
outcome_names <- c('Inhalers')

# populate outcome column
rr_df[,2] <- factor(outcome_names, levels=unique(outcome_names))

## quasipoisson
smk_est_fq <- tidy(glm(as.formula(paste0(outcome_list, 
    "~ geo_smk_pm10 + wrf_temp_county + county + 
    day + offset(log(pop))")), 
    or_ndc_county_ts, family="quasipoisson"))[2, ]

rr_fq <- c(round(exp(smk_est_fq[1,2]),5),
         round(exp(smk_est_fq[1,2]-1.96*smk_est_fq[1,3]),5),
         round(exp(smk_est_fq[1,2]+1.96*smk_est_fq[1,3]),5))

rr_df[1,3:5] <- rr_fq

## poisson
smk_est_fp <- tidy(glm(as.formula(paste0(outcome_list, 
    "~ geo_smk_pm10 + wrf_temp_county + county + 
    day + offset(log(pop))")), 
    or_ndc_county_ts, family="poisson"))[2, ]

rr_fp <- c(round(exp(smk_est_fp[1,2]),5),
         round(exp(smk_est_fp[1,2]-1.96*smk_est_fp[1,3]),5),
         round(exp(smk_est_fp[1,2]+1.96*smk_est_fp[1,3]),5))

rr_df[2,3:5] <- rr_fp

## mixed
smk_est_mp <- tidy(glmer(as.formula(paste0(outcome_list, 
    "~ geo_smk_pm10 + wrf_temp_county + (1|county) + 
    day + offset(log(pop))")), 
    or_ndc_county_ts, family="poisson"))[2, ]

rr_df[3,3:5] <- rr_mp

rr_df

plot_rr <- ggplot(rr_df, aes(x=method, y = rel_risk)) +
  geom_point() +
  geom_errorbar(aes(ymin=lower_95, ymax=upper_95), width = 0.2) +
  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
  theme(
   panel.background = element_rect(fill = 'white', colour = 'black'),
   panel.grid.major = element_blank(),
   panel.grid.minor = element_blank(),
   axis.text.x = element_text(angle = 90, vjust = .75)) +
  ylab("Relative Risk") +
  xlab("Outcome") +
  ggtitle("Relative Risk for a 10ug/m^3 increase PM2.5 estimated via GWR Smoke")

ggsave("../plot_new/ndc_rr_geo.pdf", plot = plot_rr, 
       width = 12, height = 8, units = "in")

## Distributed Lag Model

### Lag 1

Remove the missing values for GWR smoke. 
But I use GWR smoke of zip for the case crossover study.

```{r DLM, echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}
read_path8 <- paste0("../data_new/county_data/zip_to_county_new.csv")
or_zip_to_county <- read_csv(read_path8)

@@

inhaler_ndc_lag_df <- inhaler_ndc_df %>%
  left_join(or_zip_to_county, by = "ZIPCODE") %>%
  left_join(or_pop, by = "county") %>%
  mutate(day = ifelse(weekdays(date) == "Monday" |
                      weekdays(date) == "Tuesday" |
                      weekdays(date) == "Wednesday"|
                      weekdays(date) == "Thursday"|
                      weekdays(date) == "Friday", "weekday",
               ifelse(weekdays(date) == "Saturday" |
                      weekdays(date) == "Sunday", "weekend", NA))) %>%
  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10) %>%
  filter(!is.na(geo_smk_pm10_lag7_zip))

@@

which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_zip") # 159
which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_lag1_zip") # 160

inhaler_ndc_lag <- inhaler_ndc_lag_df %>%
  select(159, 160:166) # GWR-method

lag_max <- 7
n <- length(inhaler_ndc_lag[,1])

inhaler_ndc_lag <- as.matrix(inhaler_ndc_lag)

# the nrow B should equal ncol(pm)
dim(B)
dim(inhaler_ndc_lag)

inhaler_ndc_lag_B <- inhaler_ndc_lag%*%B
fit <- clogit(outcome ~ inhaler_ndc_lag_B + wrf_temp_zip + strata(personkey), inhaler_ndc_lag_df)

coef(fit)
summary(fit)

dlparms <- grep("inhaler_ndc_lag_B",names(coef(fit)))
DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms])  
DLvar <- B%*%vcov(fit)[dlparms,dlparms]%*%t(B)
DLestimate$SE <- sqrt(diag(DLvar))
DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96

cumulative <- sum(DLestimate$estimate)
cumulative
## [1] 0.009141771
cumulative_se <- sqrt(sum(DLvar))
cumulative_se
## [1] 0.000955469
cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
cumulative_CI
## [1] 0.007269052 0.011014490

## mixed model
mod <-glmer(outcome ~  inhaler_ndc_lag_B + wrf_temp_zip + day + (1|county) + offset(log(pop)),   
            inhaler_ndc_lag_df, family="poisson", control = glmerControl(optimizer = "bobyqa"))

summary(mod)
# check AIC
AIC(mod)
# output distributed lag beta parameters
dlparms <- mod@beta[2:4]
# estimate distributed lag values for each day
dl_estimates <- data.frame(estimate = B %*% dlparms)

```{r, message=FALSE, echo=FALSE, eval=FALSE}
# create fit matrix
fit_mat <- matrix(NA, nrow=4, ncol=6)
colnames(fit_mat) <- c("outcome", "exposure", "fit_mix", "df", "aic", "fit_fix")

@@

    # define basis b using natural spline function
    b <- ns(0:7, df = degree_freedom, intercept = T)

    # multiply lagged pm matrix by basis
    pm_b <- inhaler_ndc_lag %*% b

    # fit mixed model
    mod <-glmer(outcome ~  pm_b + (1|county) + offset(log(pop)), inhaler_ndc_lag_df, family="poisson", control = glmerControl(optimizer = "bobyqa"))

    # fixed mod
    fixed_mod <- glm(outcome ~ pm_b + county + offset(log(pop)), inhaler_ndc_lag_df, 
                     family = "poisson")

    AIC(mod)
    AIC(fixed_mod)
    # fill AIC
    fit_mat[h-2,5] <- round(AIC(mod),5)
    fit_mat[h-2,6] <- round(AIC(fixed_mod),5)

@@

### Compare fixed and mixed models of df 3 and 4

```{r compare mixed and fixed, message=FALSE, echo=FALSE, eval=FALSE}

## degree of 3 or 4, mixed or fixed
degree_freedom <- 3

# define basis b using natural spline function
b <- ns(0:7, df = degree_freedom, intercept = T)

# multiply lagged pm matrix by basis
pm_b <- inhaler_ndc_lag %*% b

mod <-glmer(outcome ~  pm_b + (1|county) + wrf_temp_zip + offset(log(pop)), inhaler_ndc_lag_df, family="poisson",
            control = glmerControl(optimizer = "bobyqa"))

dlparms <- mod@beta[2:(1+degree_freedom)]
# estimate distributed lag values for each day
dl_estimates <- data.frame(estimate = b %*% dlparms)

# covariance matrix for knots (need to convert to matrix object)
cov_mat <- as.matrix(vcov(mod)[2:(1+degree_freedom),2:(1+degree_freedom)])

# estimate variance of splines
dl_var <- b%*%cov_mat%*%t(b)

# plot 
p3_mix <- ggplot(inhaler_dl_rr, aes(x=lag, y=estimate, 
                              ymin=lower_bound, ymax=upper_bound)) +
  geom_ribbon(alpha = 0.5) +
  geom_line(size=2)

fixed_mod <- glm(outcome ~ pm_b + county + wrf_temp_zip + offset(log(pop)), inhaler_ndc_lag_df, 
                     family = "poisson")

dlparms <- fixed_mod$coefficients[2:4]
# estimate distributed lag values for each day
dl_estimates <- data.frame(estimate = b %*% dlparms)

# covariance matrix for knots (need to convert to matrix object)
cov_mat <- as.matrix(vcov(fixed_mod)[2:4,2:4])

# estimate variance of splines
dl_var <- b%*%cov_mat%*%t(b)

# calculate lower and upper bounds
dl_estimates$lower_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(1-0.975,
  df=df.residual(fixed_mod))  
dl_estimates$upper_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(0.975,
  df=df.residual(fixed_mod))

# estimate cumulative CI
cumulative_ci <- cumulative+cumulative_se*qt(c(1-0.975,0.975), 
                                             df=df.residual(fixed_mod))
cumulative_ci

# plot 
p3_fix <- ggplot(inhaler_dl_rr, aes(x=lag, y=estimate, 
                              ymin=lower_bound, ymax=upper_bound)) +
  geom_ribbon(alpha = 0.5) +
  geom_line(size=2)

p3_fix
p3_mix

## df = 4
degree_freedom <- 4

# define basis b using natural spline function
b <- ns(0:7, df = degree_freedom, intercept = T)

# multiply lagged pm matrix by basis
pm_b <- inhaler_ndc_lag %*% b

mod <-glmer(outcome ~  pm_b + (1|county) + wrf_temp_zip + offset(log(pop)), inhaler_ndc_lag_df, family="poisson",
            control = glmerControl(optimizer = "bobyqa"))

dlparms <- mod@beta[2:5]
# estimate distributed lag values for each day
dl_estimates <- data.frame(estimate = b %*% dlparms)

# covariance matrix for knots (need to convert to matrix object)
cov_mat <- as.matrix(vcov(mod)[2:5,2:5])

# estimate variance of splines
dl_var <- b%*%cov_mat%*%t(b)

# plot 
p4_mix <- ggplot(inhaler_dl_rr, aes(x=lag, y=estimate, 
                              ymin=lower_bound, ymax=upper_bound)) +
  geom_ribbon(alpha = 0.5) +
  geom_line(size=2)

fixed_mod <- glm(outcome ~ pm_b + county + wrf_temp_zip + offset(log(pop)), inhaler_ndc_lag_df, 
                     family = "poisson")

dlparms <- fixed_mod$coefficients[2:5]
# estimate distributed lag values for each day
dl_estimates <- data.frame(estimate = b %*% dlparms)

# covariance matrix for knots (need to convert to matrix object)
cov_mat <- as.matrix(vcov(fixed_mod)[2:5,2:5])

# estimate variance of splines
dl_var <- b%*%cov_mat%*%t(b)

# calculate lower and upper bounds
dl_estimates$lower_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(1-0.975,
  df=df.residual(fixed_mod))  
dl_estimates$upper_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(0.975,
  df=df.residual(fixed_mod))

# estimate cumulative CI
cumulative_ci <- cumulative+cumulative_se*qt(c(1-0.975,0.975), 
                                             df=df.residual(fixed_mod))
cumulative_ci

# plot 
p4_fix <- ggplot(inhaler_dl_rr, aes(x=lag, y=estimate, 
                              ymin=lower_bound, ymax=upper_bound)) +
  geom_ribbon(alpha = 0.5) +
  geom_line(size=2)

p4_fix
p4_mix

ggsave("../plot_new/ndc_p3_mix.pdf", plot = p3_mix, 
       width = 12, height = 8, units = "in")
ggsave("../plot_new/ndc_p4_mix.pdf", plot = p4_mix, 
       width = 12, height = 8, units = "in")

ggsave("../plot_new/ndc_p3_fix.pdf", plot = p3_fix, 
       width = 12, height = 8, units = "in")
ggsave("../plot_new/ndc_p4_fix.pdf", plot = p4_fix, 
       width = 12, height = 8, units = "in")

@@

Note I get some warnings (including max grade convergence issues), slightly different answers in best fit, and it takes about 18 minutes to run. Consider running in parallel.
I suspect the convergence issues are likely due to the same state/counties with small sample size and few outcome observations and lack of variability between exposure/outcome.
Going to fit what I have now, take a walk, and see what comes out. First thing to do is find cumulative effect, then lagged day effects.

```{r, message=FALSE, echo=FALSE, include=FALSE, eval=FALSE}
# create a dataframe that contains cumulative estimates for each outcome
cumulative_df <- as_tibble(matrix(NA, nrow=0, ncol=4))
colnames(cumulative_df) <- c("outcome", "rel_risk", "lower_95", "upper_95")

# create a dataframe that contains the distributed lag estimates for each outcome
dist_lag_df <- as_tibble(matrix(NA,  nrow = 0, ncol = 5))
colnames(dist_lag_df) <- c("outcome", "lag", "rel_risk", 
                                  "lower_95", "upper_95")

# loop through best fit df

    # create empty matrix two empty matrices to fill
    cumul_val <- as_tibble(matrix(NA, nrow=1, ncol=4))
    colnames(cumul_val) <- c("outcome", "rel_risk", "lower_95", "upper_95")

    dl_val <- as_tibble(matrix(NA, nrow=8, ncol=5))
    colnames(dl_val) <- c("outcome", "lag", "rel_risk", 
                                  "lower_95", "upper_95")

    # easiest to set out the parameters from the start
    outcome <- "inhalers"
    degree_freedom <- as.numeric(lag_spline_best_fit[4])

    # define basis b using natural spline function
    b <- ns(0:7, df = degree_freedom, intercept = T)

    # multiply lagged pm matrix by basis
    pm_b <- inhaler_ndc_lag %*% b

    # fit mixed model
     mod <-glmer(outcome ~  inhaler_ndc_lag_B + wrf_temp_zip + day + (1|county) + offset(log(pop)), inhaler_ndc_lag_df, family="poisson", control = glmerControl(optimizer = "bobyqa"))

    mod
    # output distributed lag beta parameters; this will change based on DF
    dlparms <- mod@beta[2:(1+degree_freedom)]
    # estimate distributed lag values for each day
    dl_estimates <- data.frame(estimate = b %*% dlparms)

    # covariance matrix for knots (need to convert to matrix object)
    cov_mat <- as.matrix(vcov(mod)[2:(1+degree_freedom),2:(1+degree_freedom)]) 
    # estimate variance of splines
    dl_var <- b%*%cov_mat%*%t(b)
    # calculate standard error for each lag value
    dl_estimates$stderr <- sqrt(diag(dl_var))
    # calculate lower and upper bounds
    dl_estimates$lower_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(1-0.975,
      df=df.residual(mod))  
    dl_estimates$upper_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(0.975,
      df=df.residual(mod))

    # estimate cumulative effect ----
    cumulative <- sum(dl_estimates$estimate)
    # estimate cumulative effect stnd error
    cumulative_se <- sqrt(sum(dl_var))
    # estimate cumulative CI
    cumulative_ci <- cumulative+cumulative_se*qt(c(1-0.975,0.975), 
                                                 df=df.residual(mod))
    # fill tibbles and bind in to larger dataframes ----
    # outcome
    cumul_val[1,1] <- outcome
    # cumulative rel risk
    cumul_val[1,2] <- exp(cumulative)
    # cumulative 95% CI
    cumul_val[1,3:4] <- exp(cumulative_ci)
    # bind to cumulative dataframe
    cumulative_df <- rbind(cumulative_df, cumul_val)

    # lag vals (this step is redundant)
    dl_val[,1] <- outcome
    dl_val[,2] <- 0:7
    dl_val[,3] <- exp(dl_estimates[,1])
    dl_val[,4] <- exp(dl_estimates[,3])
    dl_val[,5] <- exp(dl_estimates[,4])

    # bind in with overall df
    dist_lag_df <- rbind(dist_lag_df, dl_val)

# print table of cumulative effects
knitr::kable(cumulative_df, caption = paste0("Cumulative Effect for a 10 ug/m^3 increase in GWR smoke lagged days 0-7 mixed model approach"))

# plot distributed lag effects
ggplot(dist_lag_df, aes(x=lag, y=rel_risk)) +
  geom_line(colour = "blue") +
  geom_ribbon(aes(ymin = lower_95, ymax = upper_95), fill = "blue", alpha = 0.3) + 
  scale_y_continuous(limits = c(0.98, 1.06)) +
  scale_x_continuous(breaks = c(seq(0, 7,by=1))) +
  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
  facet_wrap(~ outcome) +
  ggtitle(paste0("Distributed lag association between a 10 ug/m^3",
                 " increase in GWR smoke PM2.5 and cardiopulmonary outcomes")) +
  ylab("Relative Risk for a 10 ug/m^3 increase in GWR smoke (fixed)") +
  xlab("Lagged Days") +
  theme_bw()

@@

## Distributed Lag -- Casecrossover Analysis

### 1. For Overall Plot (Unconstrained lag model)

@@

### 2. Distributed lag for casecrossover

```{r DLM for casecrossover, echo = FALSE, message=FALSE, warning=FALSE}
read_path8 <- paste0("../data_new/county_data/zip_to_county_new.csv")
or_zip_to_county <- read_csv(read_path8)

# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
dlparms <- c(1,2,3,4,5,6)
DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms]) 
DLvar <- D%*%vcov(fit)[dlparms,dlparms]%*%t(D)
DLestimate$SE <- sqrt(diag(DLvar))
DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96

```{r DLM for casecrossover temp, echo = FALSE, message=FALSE, warning=FALSE}
read_path8 <- paste0("../data_new/county_data/zip_to_county_new.csv")
or_zip_to_county <- read_csv(read_path8)

### 3. Unconstrained lag for time series

```{r unconstrained ts lag, message=FALSE, echo=FALSE}
or_ndc_county_ts_df <- or_ndc_county_ts %>%
  select(1:4, pop, rate_per_100k, day, geo_smk_pm_county,
         geo_smk_pm_lag1_county:geo_smk_pm_lag7_county, wrf_temp_county,
         wrf_temp_lag1_county:wrf_temp_lag7_county) %>%
  mutate(geo_smk_pm10_county = geo_smk_pm_county/10,
         geo_smk_pm10_lag1_county =geo_smk_pm_lag1_county/10,
         geo_smk_pm10_lag2_county =geo_smk_pm_lag2_county/10,
         geo_smk_pm10_lag3_county =geo_smk_pm_lag3_county/10,
         geo_smk_pm10_lag4_county =geo_smk_pm_lag4_county/10,
         geo_smk_pm10_lag5_county =geo_smk_pm_lag5_county/10,
         geo_smk_pm10_lag6_county =geo_smk_pm_lag6_county/10,
         geo_smk_pm10_lag7_county =geo_smk_pm_lag7_county/10) 
  # filter(!is.na(geo_smk_pm10_lag7_zip))

@@

### 4. Distributed lag for time series

```{r DLM ts, echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}

which(colnames(or_ndc_county_ts_df)=="wrf_temp_lag1_county") # 17
which(colnames(or_ndc_county_ts_df)=="geo_smk_pm10_lag1_county") # 25

@@

## mixed model
mod <-glmer(outcome ~  geo_ts_lag_B + tem_ts_lag_B + day + (1|county) + offset(log(pop)),   
            or_ndc_county_ts_df, family="poisson", control = glmerControl(optimizer = "bobyqa"))

>>>>

### Time-Stratified Case-Crossover (Population-Weighted PM~2.5~ Assigned by Zip Code)
*referents same day of week within fire season May - September*

@@

### Time stratified for time series (Population-Weighted PM~2.5~ Assigned by County)

I also make time stratified time series data set. The original data set only contains the zip code, but for time series, we want to research on the county level. Therefore, I assign each zip code with one county which I found online. There are 488 zipcodes in total (4 are missing but I use google to find it).

@@

## 1. smoke exposure time series and map of smoke days in Oregon

print(plot_county)
# print(plot_county_line)
# save figure
ggsave("../plot_new/ndc_county_rate_ts.pdf", plot = plot_county, 
       width = 12, height = 8, units = "in")
ggsave("../plot_new/ndc_county_rate_ts_line.pdf", plot = plot_county_line, 
       width = 12, height = 8, units = "in")

plot(plot_health_smk)
# plot(plot_county_health_smk)

## 2. descriptive table of total ndc inhalhers; and age and sex category; Plots

@@

## 3. Distributed Lag Model

Remove the missing values for GWR smoke. 
But I only use GWR smoke of zip for the case crossover study.

Choose df = 5 because it has the least AIC.

```{r, message=FALSE, echo=FALSE, warning=FALSE}
read_path8 <- paste0("../data_new/county_data/zip_to_county_new.csv")
or_zip_to_county <- read_csv(read_path8)

@@

@@

# which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_zip") # 159
# which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_lag1_zip") # 160
# which(colnames(inhaler_ndc_lag_df)=="wrf_temp_zip") # 93
# which(colnames(inhaler_ndc_lag_df)=="wrf_temp_lag1_zip") # 150

# create fit matrix
fit_mat <- matrix(NA, nrow=4, ncol=5)
colnames(fit_mat) <- c("outcome", "exposure", "fit", "df", "aic")

@@

    # define basis b using natural spline function
    B <- ns(0:7, df = degree_freedom, intercept = T)

    # multiply lagged pm matrix by basis
    inhaler_geo_lag_B <- inhaler_geo_lag%*%B
    inhaler_temp_lag_B <- inhaler_temp_lag%*%B

    AIC(fit)

    # fill AIC
    fit_mat[h-2,5] <- round(AIC(fit),5)

@@

### Distributed lag 0-7 for casecrossover

```{r DLM for casecrossover, echo = FALSE, message=FALSE, warning=FALSE}
inhaler_geo_lag <- inhaler_ndc_lag_df %>%
  select(159, 160:166) # GWR-method

B <- ns(0:lag_max, df =5, intercept=TRUE)

# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
dlparms <- c(1,2,3,4,5,6,7,8,9,10)
DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms]) 
DLvar <- D%*%vcov(fit)[dlparms,dlparms]%*%t(D)
DLestimate$SE <- sqrt(diag(DLvar))
DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96

@@

```{r DLM for casecrossover temp, echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}
read_path8 <- paste0("../data_new/county_data/zip_to_county_new.csv")
or_zip_to_county <- read_csv(read_path8)

@@

### Distributed lag -7 to 0 for casecrossover

```{r time stratified zip for casecrossover lag -7 to 0 prepare, eval=FALSE, echo=FALSE}
### Time stratified ------------------------------------------------------------
# read in zipcode level populatoin-weighted pm
read_path5 <- paste0('C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire_new/data/Oregon_PM/zip_pm_to_merge_with_acap.csv')
         wrf_temp_lag7 = lag(wrf_temp, 7, order_by = ZIPCODE),
         # geo lead
         geo_smk_pm_lag_m1 = lead(geo_smk_pm, 1, order_by = ZIPCODE),
         geo_smk_pm_lag_m2 = lead(geo_smk_pm, 2, order_by = ZIPCODE),
         geo_smk_pm_lag_m3 = lead(geo_smk_pm, 3, order_by = ZIPCODE),
         geo_smk_pm_lag_m4 = lead(geo_smk_pm, 4, order_by = ZIPCODE),
         geo_smk_pm_lag_m5 = lead(geo_smk_pm, 5, order_by = ZIPCODE),
         geo_smk_pm_lag_m6 = lead(geo_smk_pm, 6, order_by = ZIPCODE),
         geo_smk_pm_lag_m7 = lead(geo_smk_pm, 7, order_by = ZIPCODE),
         # temp lead
         wrf_temp_lag_m1 = lead(wrf_temp, 1, order_by = ZIPCODE),
         wrf_temp_lag_m2 = lead(wrf_temp, 2, order_by = ZIPCODE),
         wrf_temp_lag_m3 = lead(wrf_temp, 3, order_by = ZIPCODE),
         wrf_temp_lag_m4 = lead(wrf_temp, 4, order_by = ZIPCODE),
         wrf_temp_lag_m5 = lead(wrf_temp, 5, order_by = ZIPCODE), 
         wrf_temp_lag_m6 = lead(wrf_temp, 6, order_by = ZIPCODE),
         wrf_temp_lag_m7 = lead(wrf_temp, 7, order_by = ZIPCODE)) %>% 

write_path4 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover_zip_lag15.csv")
write_csv(outcome_casecross, write_path4)

```{r time stratified county for ts lag -7 to 0 prepare, echo=FALSE, include=FALSE, eval=FALSE}
read_path7 <- paste0("../data_new/medication/oregon_asthma_ndc.csv")
or_ndc_claim <- read_csv(read_path7)
         wrf_temp_lag7 = lag(wrf_temp, 7, order_by = county),
          # geo lead
         geo_smk_pm_lag_m1 = lead(geo_smk_pm, 1, order_by = county),
         geo_smk_pm_lag_m2 = lead(geo_smk_pm, 2, order_by = county),
         geo_smk_pm_lag_m3 = lead(geo_smk_pm, 3, order_by = county),
         geo_smk_pm_lag_m4 = lead(geo_smk_pm, 4, order_by = county),
         geo_smk_pm_lag_m5 = lead(geo_smk_pm, 5, order_by = county),
         geo_smk_pm_lag_m6 = lead(geo_smk_pm, 6, order_by = county),
         geo_smk_pm_lag_m7 = lead(geo_smk_pm, 7, order_by = county),
         # temp lead
         wrf_temp_lag_m1 = lead(wrf_temp, 1, order_by = county),
         wrf_temp_lag_m2 = lead(wrf_temp, 2, order_by = county),
         wrf_temp_lag_m3 = lead(wrf_temp, 3, order_by = county),
         wrf_temp_lag_m4 = lead(wrf_temp, 4, order_by = county),
         wrf_temp_lag_m5 = lead(wrf_temp, 5, order_by = county), 
         wrf_temp_lag_m6 = lead(wrf_temp, 6, order_by = county),
         wrf_temp_lag_m7 = lead(wrf_temp, 7, order_by = county)) %>% 

write_path5 <- paste0("../data_new/medication/or_ndc_county_time_series_lag15.csv")
write_csv(or_ndc_county_ts, write_path5)

```{r time stratified zip for casecrossover lag -7 to 0 df, echo=FALSE, include=FALSE}
read_path10 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover_zip_lag15.csv")
inhaler_ndc_lag15 <- read_csv(read_path10)
# asthma_ndc <- fread(read_path6, showProgress = T) # faster, but seems error in overall plot analysis

# dataframe to loop through
inhaler_ndc_lag15_df <- data.frame(inhaler_ndc_lag15)

inhaler_ndc_lag_15_df <- inhaler_ndc_lag15_df %>%
  mutate(geo_smk_pm10_lag_m1_zip =geo_smk_pm_lag_m1_zip/10,
         geo_smk_pm10_lag_m2_zip =geo_smk_pm_lag_m2_zip/10,
         geo_smk_pm10_lag_m3_zip =geo_smk_pm_lag_m3_zip/10,
         geo_smk_pm10_lag_m4_zip =geo_smk_pm_lag_m4_zip/10,
         geo_smk_pm10_lag_m5_zip =geo_smk_pm_lag_m5_zip/10,
         geo_smk_pm10_lag_m6_zip =geo_smk_pm_lag_m6_zip/10,
         geo_smk_pm10_lag_m7_zip =geo_smk_pm_lag_m7_zip/10) %>%
  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)

# which(colnames(inhaler_ndc_lag_15_df)=="geo_smk_pm10_zip") # 178
# which(colnames(inhaler_ndc_lag_15_df)=="geo_smk_pm10_lag_m1_zip") # 171
# which(colnames(inhaler_ndc_lag_15_df)=="geo_smk_pm10_lag1_zip") # 179

# which(colnames(inhaler_ndc_lag_15_df)=="wrf_temp_zip") # 93
# which(colnames(inhaler_ndc_lag_15_df)=="wrf_temp_lag1_zip") # 150
# which(colnames(inhaler_ndc_lag_15_df)=="wrf_temp_lag_m1_zip") # 164

inhaler_geo_lag <- inhaler_ndc_lag_15_df %>%
  select(177:171, 178) # GWR-method

inhaler_temp_lag <- inhaler_ndc_lag_15_df %>%
  select(170:164, 93) # temp

# create fit matrix
fit_mat <- matrix(NA, nrow=4, ncol=5)
colnames(fit_mat) <- c("outcome", "exposure", "fit", "df", "aic")

    # define basis b using natural spline function
    B <- ns(-7:0, df = degree_freedom, intercept = T)

    fit <- clogit(outcome ~ inhaler_geo_lag_B + inhaler_temp_lag_B + strata(personkey), 
              inhaler_ndc_lag_15_df)

    AIC(fit)

    # fill AIC
    fit_mat[h-2,5] <- round(AIC(fit),5)

# kable
knitr::kable(lag_spline_best_fit, caption = paste0("Distributed lag spline",
  "degree of freedom best fit by AIC for mixed model approach"))

```{r DLM for casecrossover lag -7 to 0, echo = FALSE, message=FALSE, warning=FALSE}
inhaler_geo_lag <- inhaler_ndc_lag_15_df %>%
  select(177:171, 178) # GWR-method

inhaler_temp_lag <- inhaler_ndc_lag_15_df %>%
  select(170:164, 93) # temp

lag_max <- 0
n1 <- length(inhaler_geo_lag[,1])
n2 <- length(inhaler_temp_lag[,1])

B <- ns(-7:lag_max, df =5, intercept=TRUE)

fit <- clogit(outcome ~ inhaler_geo_lag_B + inhaler_temp_lag_B + strata(personkey), 
              inhaler_ndc_lag_15_df)

# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
dlparms <- c(1,2,3,4,5,6,7,8,9,10)
DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms]) 
DLvar <- D%*%vcov(fit)[dlparms,dlparms]%*%t(D)
DLestimate$SE <- sqrt(diag(DLvar))
DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96

DLestimate$lag <- -7:lag_max
p <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper))
p <- p + geom_ribbon(alpha=.5) + geom_line(size=2)
p <- p + geom_line(aes(x=lag), linetype=2, size=2)
p <- p + theme_bw()
p

## Supplement
### 1. For Overall Plot (Unconstrained lag model)

@@

### 2. Unconstrained lag for time series

```{r unconstrained ts lag, message=FALSE, echo=FALSE, warning=FALSE}
or_ndc_county_ts_df <- or_ndc_county_ts %>%
  select(1:4, pop, rate_per_100k, day, geo_smk_pm_county,
         geo_smk_pm_lag1_county:geo_smk_pm_lag7_county, wrf_temp_county,
         wrf_temp_lag1_county:wrf_temp_lag7_county) %>%
  mutate(geo_smk_pm10_county = geo_smk_pm_county/10,
         geo_smk_pm10_lag1_county =geo_smk_pm_lag1_county/10,
         geo_smk_pm10_lag2_county =geo_smk_pm_lag2_county/10,
         geo_smk_pm10_lag3_county =geo_smk_pm_lag3_county/10,
         geo_smk_pm10_lag4_county =geo_smk_pm_lag4_county/10,
         geo_smk_pm10_lag5_county =geo_smk_pm_lag5_county/10,
         geo_smk_pm10_lag6_county =geo_smk_pm_lag6_county/10,
         geo_smk_pm10_lag7_county =geo_smk_pm_lag7_county/10) 
  # filter(!is.na(geo_smk_pm10_lag7_zip))

@@

### 3. Distributed lag check for time series

```{r DLM ts, echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}
which(colnames(or_ndc_county_ts_df)=="wrf_temp_lag1_county") # 17
which(colnames(or_ndc_county_ts_df)=="geo_smk_pm10_lag1_county") # 25

@@

## mixed model
mod <-glmer(n_obs ~  geo_ts_lag_B + tem_ts_lag_B + day + (1|county) + offset(log(pop)),   
            or_ndc_county_ts_df, family="poisson", control = glmerControl(optimizer = "bobyqa"))

summary(mod)
# check AIC
AIC(mod)
